### 2021-10-08 manuscript/cover_letter.tex
Updates from Overleaf

\documentclass[11pt]{letter}

\usepackage[bitstream-charter]{mathdesign}
\usepackage{microtype}
\usepackage{graphicx}
\usepackage[letterpaper, margin=0.9in, top=0.4in, bottom=0.4in]{geometry}

\longindentation=30pt

\address{Anna {\AA}kesson \\ Link\"oping University \\ Division of Theoretical Biology, Dept.~IFM \\ SE-58183 Link\"oping, Sweden}

\date{October XX, 2021}

\begin{document}

\begin{letter}{
\vspace{-2.5cm}

Dr.~XX\\
Editor-in-Chief\\
Proceedings of the Royal Society B} % Name/title of the addressee

\opening{Dear Dr.~XX,}

Please find enclosed the manuscript ``Does motif participation provide unique information about species risk of extinction?'' by Anna {\AA}kesson et al., which we would like to be considered for publication in \emph{Proceedings of the Royal Society B} as an Article.

Alyssa Cirtwill has earlier been approached by the Proceedings B Preprint Editorial Team, considering another paper regarding motifs.

While there is pressing need to understand ...
Here we study the effects of these processes ... 
Several key results emerge ...

First ...

Second ...

Third ...

We reach our conclusions by ...

Due to the relevance of our results to understanding ..., for which \emph{ Proceedings of the Royal Society B} would be an ideal platform.

We thank you for your time and attention dedicated to our work.

\vspace{2\parskip}

\hspace{4\parskip} Sincerely,

\hspace{4\parskip} Anna {\AA}kesson, Alyssa R. Cirtwill, Kate Wootton, and Anna Ekl\"of

\end{letter} 
\end{document}