### 2021-10-14 manuscript/cover_letter.tex
Updates from Overleaf

\documentclass[11pt]{letter}

\usepackage[bitstream-charter]{mathdesign}
\usepackage{microtype}
\usepackage{graphicx}
\usepackage[letterpaper, margin=0.9in, top=0.4in, bottom=0.4in]{geometry}

\longindentation=30pt

\address{Anna {\AA}kesson \\ Link\"oping University \\ Division of Theoretical Biology, Dept.~IFM \\ SE-58183 Link\"oping, Sweden}

\date{October XX, 2021}

\begin{document}

\begin{letter}{
\vspace{-2.5cm}

Dr.~XX\\
Editor-in-Chief\\
Proceedings of the Royal Society B} % Name/title of the addressee

\opening{Dear Dr.~XX,}

Please find enclosed the manuscript ``Does motif participation provide unique information about species risk of extinction?'' by Anna {\AA}kesson et al., which we would like to be considered for publication in \emph{Proceedings of the Royal Society B} as an Article.

Alyssa Cirtwill was asked by the Proceedings B Preprint Editorial Team to consider publishing a thematically similar paper in this journal. That manuscript is currently in revisions elsewhere, but we consider that this manuscript will also be a good topical fit. Both manuscripts deal with the relationship between species' motif roles and risk of extinction, although in different modelling frameworks and using different types of disturbance.

While there is pressing need to understand ...
Here we study the effects of these processes ... 
Several key results emerge ...

First ...

Second ...

Third ...

We reach our conclusions by ...

Due to the relevance of our results to understanding ..., for which \emph{ Proceedings of the Royal Society B} would be an ideal platform.

We thank you for your time and attention dedicated to our work.

%The cover letter is used to tell the handling editor why you have submitted to this journal, and provide any additional information that might not be available within the manuscript, for example setting the paper in context within the field, alerting the editor to related papers in the field, or explaining choices of preferred or non-preferred reviewers.

\vspace{2\parskip}

\hspace{4\parskip} Sincerely,

\hspace{4\parskip} Anna {\AA}kesson, Alyssa R. Cirtwill, Kate Wootton, and Anna Ekl\"of

\end{letter} 
\end{document}